### 2014-03-20 presentation/presentation.tex
Related work

\documentclass{beamer}

\usepackage{acronym}
\usepackage[backend=bibtex]{biblatex}
\usepackage{default}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{lmodern}

\input{../common/acronyms}
\bibliography{../common/bibliography.bib}

\setbeamertemplate{bibliography item}{}

\title{Concurrent Priority Queues}
\subtitle{Seminar in Algorithms, 2013W}
\author{Jakob Gruber, 0203440}
\date{\today}

\begin{document}

\maketitle

\begin{frame}{Outline}
\begin{minipage}[t][10em][t]{\linewidth}
\tableofcontents
\end{minipage}
\end{frame}

% --------------------------------------------------------------------------------------------------
\section{Introduction}
% --------------------------------------------------------------------------------------------------

\begin{frame}{Hello World}
Hello
\end{frame}

% --------------------------------------------------------------------------------------------------
\section{Related Work}
% --------------------------------------------------------------------------------------------------

\begin{frame}{Related Work}
\begin{itemize}
\item Non-standard synchronization primitives
    \begin{itemize}
    \item \citeauthor{liu2012lock}: Array-based PQ with \lstinline|ExtractMany|
    \item \citeauthor{israeli1993efficient}: Wait-free PQ
    \end{itemize}

\item Bounded range priorities
    \begin{itemize}
    \item \citeauthor{shavit1999scalable}: Combining funnels \& bins
    \end{itemize}

\item PQs in distributed memory systems
    \begin{itemize}
    \item \citeauthor{sanders1998randomized}: Local PQs, \lstinline|Insert| at random processor
    \end{itemize}

\item Relaxed data structures
    \begin{itemize}
    \item \citeauthor{kirsch2012fast}: k-FIFO queues
    \end{itemize}

\end{itemize}
\end{frame}

% --------------------------------------------------------------------------------------------------
\section{Concepts and Definitions}
% --------------------------------------------------------------------------------------------------

% --------------------------------------------------------------------------------------------------
\section{Fine-grained Locking Heaps}
% --------------------------------------------------------------------------------------------------

% --------------------------------------------------------------------------------------------------
\section{Lock-free Priority Queues}
% --------------------------------------------------------------------------------------------------

% --------------------------------------------------------------------------------------------------
\section{Relaxed Priority Queues}
% --------------------------------------------------------------------------------------------------

% --------------------------------------------------------------------------------------------------
\section{Conclusion}
% --------------------------------------------------------------------------------------------------

% --------------------------------------------------------------------------------------------------
\section{References}
% --------------------------------------------------------------------------------------------------

\begin{frame}[allowframebreaks]{References}
\printbibliography
\end{frame}

\end{document}